Stellar Bars in Isolated Gas-Rich Galaxies Do Not Slow Down

### 2022-04-04 paper/ms.tex
Add dark matter halo
Added the phrase "dark matter" to some references of the halo where it might be unclear to the reader

\begin{figure*}[h]%
\centering
\includegraphics[width=0.9\textwidth]{fig/fig3.pdf}
\caption{The wake excited in the dark matter halo in the N-body case
(\textit{left panel}) and SMUGGLE case (\textit{right panel}) after
$2.6\,\textrm{Gyr}$. The \textit{left} and \textit{right panels} show a
surface density projection in the $x$-$y$ plane of the dark matter halo after
an axisymmetric average has been subracted. The solid line indicates the
direction of the bar while the dashed line indicates the direction of the halo
wake (both measured by taking the second Fourier component within a sphere of
all material within a radius of $X\,\textrm{kpc}$). The \textit{center panel}
shows the time evolution of the angle difference between the bar and the halo
wake, as measured from the second Fourier component. After the first Gyr, the
angle difference in the SMUGGLE case is smaller than in the N-body case by
about a factor of two, reflecting how the halo in the SMUGGLE case is unable
to exert as negative a torque on the bar as in the N-body
case.}\label{fig:wake}
\end{figure*}

The presence of the gas can arrest the process by which the halo wake forms.
However, this does not explain why the pattern speed in the SMUGGLE case is
nearly constant over several Gyr. Naively, it would be a coincidence that the
bar pattern speed remains constant in the SMUGGLE case, resulting from a
chance cancellation of the halo and gas torques. However, a constant pattern
speed in the presence of gas has been observed in a few simulations of barred
galaxies with gas.\cite{1993AA...268...65F, 2010ApJ...719.1470V}

We propose that an equilibrium mechanism is responsible for the pattern speed
remaining approximately constant. In this scenario, a torque must oppose
changes in the pattern speed so that when the bar speeds up, a negative torque
will slow it down and when the bar slows down, a positive torque will speed it
up. It is simple to explain the first of these - when the pattern speed
increases, the location of resonances will shift to regions of the halo phase
space where no wake has been excited yet (e.g., the corotation radius will
shrink).

@@

We have therefore shown evidence for an equilibrium mechanism at play which
keeps the pattern speed of the bar constant, resulting from the complex
interplay between the halo and the gas phase.

@@

After generating the gaseous disk in this way, we stitched the gas disk
together with the GALAKOS N-body disk (and bulge and halo) after the GALAKOS
disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The purpose of allowing
the GALAKOS to evolve first for a short period of time is to allow for the bar
to form unimpacted by the presence of the gas (which would normally disrupt
the formation of the bar). We made one additional modification when stitching
the gas disk together with the N-body disk - we created a vacuum within the
central $X$ pc. This vacuum guards against an initial dramatic infall of gas
within the bar region, which we found to destroy the bar.

>>>>

\begin{figure*}[h]%
\centering
\includegraphics[width=0.9\textwidth]{fig/fig3.pdf}
\caption{The wake excited in the dark matter halo in the N-body case
(\textit{left panel}) and SMUGGLE case (\textit{right panel}) after
$2.6\,\textrm{Gyr}$. The \textit{left} and \textit{right panels} show a
surface density projection in the $x$-$y$ plane of the dark matter halo after
an axisymmetric average has been subracted. The solid line indicates the
direction of the bar while the dashed line indicates the direction of the halo
wake (both measured by taking the second Fourier component within a sphere of
all material within a radius of $X\,\textrm{kpc}$). The \textit{center panel}
shows the time evolution of the angle difference between the bar and the halo
wake, as measured from the second Fourier component. After the first Gyr, the
angle difference in the SMUGGLE case is smaller than in the N-body case by
about a factor of two, reflecting how the dark matter halo in the SMUGGLE case
is unable to exert as negative a torque on the bar as in the N-body
case.}\label{fig:wake}
\end{figure*}

The presence of the gas can arrest the process by which the dark matter halo
wake forms. However, this does not explain why the pattern speed in the
SMUGGLE case is nearly constant over several Gyr. Naively, it would be a
coincidence that the bar pattern speed remains constant in the SMUGGLE case,
resulting from a chance cancellation of the halo and gas torques. However, a
constant pattern speed in the presence of gas has been observed in a few
simulations of barred galaxies with gas.\cite{1993AA...268...65F,
2010ApJ...719.1470V}

We propose that an equilibrium mechanism is responsible for the pattern speed
remaining approximately constant. In this scenario, a torque must oppose
changes in the pattern speed so that when the bar speeds up, a negative torque
will slow it down and when the bar slows down, a positive torque will speed it
up. It is simple to explain the first of these - when the pattern speed
increases, the location of resonances will shift to regions of the dark matter
halo phase space where no wake has been excited yet (e.g., the corotation
radius will shrink).

@@

We have therefore shown evidence for an equilibrium mechanism at play which
keeps the pattern speed of the bar constant, resulting from the complex
interplay between the dark matter halo and the gas phase.

@@

After generating the gaseous disk in this way, we stitched the gas disk
together with the GALAKOS N-body disk (and bulge and dark matter halo) after
the GALAKOS disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The
purpose of allowing the GALAKOS to evolve first for a short period of time is
to allow for the bar to form unimpacted by the presence of the gas (which
would normally disrupt the formation of the bar). We made one additional
modification when stitching the gas disk together with the N-body disk - we
created a vacuum within the central $X$ pc. This vacuum guards against an
initial dramatic infall of gas within the bar region, which we found to
destroy the bar.
